---
title: "Analyzing Aggravated Burglaries in Davidson County"
output: html_notebook
---

**Part 1 - Data Preparation**

You've been provided three datasets for this project:
  - burglaries_2023.csv: Contains data on the aggravated burglary incidents in Davidson County. This was obtained from https://data.nashville.gov/Police/Metro-Nashville-Police-Department-Incidents/2u6v-ujjs.
  - census.csv: Census tract level data on population and median income. This was obtained from the US Census American Community Survey.
  - DC: A shapefile containing Davidson County census tracts

Perform a spatial join to determine the census tract in which each burglary occurred. Hint: You may want to make use of the st_as_sf function in order to convert the burglaries data into an sf object.

After performing the spatial join, merge in the census data. Note: Make sure that the final dataset contains all census tracts.

```{r}
library('tidyverse')
library(sf)
```

1. Read the three datasets. 

```{r}
burglaries_2023 <- read_csv('../data/burglaries_2023.csv')
```
```{r}
census <- read_csv('../data/census.csv')
```

```{r}
davidson_tracts <- st_read('../data/DC/DC.shp')
```

```{r}
davidson_tracts |> 
  ggplot() + 
  geom_sf()
```

Note: Remove the two events that occurred beyond Davidson County (latitude < 34.5, longitude > 92). 

```{r}
burglaries_2023 <- burglaries_2023 |> 
  filter(!(latitude < 34.5))
```

2. Perform a spatial join to determine the census tract in which each burglary occurred.But first, use the st_as_sf function in order to convert the burglaries data into an sf object.

```{r}
burglaries_geo <- st_as_sf(burglaries_2023 |> drop_na(latitude), 
                           coords = c('longitude', 'latitude'),
                           crs = st_crs(davidson_tracts)
                           )
```

Plot the new burglaries sf object. 
```{r}
burglaries_geo |> 
  ggplot() + 
  geom_sf()
```

Plot the Davidson County tracts with the burglary (sf) data. 
```{r}
davidson_tracts |> 
  ggplot() + 
  geom_sf() + 
  geom_sf(data = burglaries_geo, size = 0.2)
```

Perform a spatial join using st_join function. 
```{r}
burglaries_tracts <- st_join(burglaries_geo, davidson_tracts, join = st_within)
```

3. After performing the spatial join, merge in the census data. Note: Make sure that the final dataset contains all census tracts.

Note: Before merging, clean up census df. Rename 'tract' column in census df to 'TRACTCE'. Remove other columns that currently exist in burglaries_tracts df. 

```{r}
census <- census |> 
  select(TRACTCE = tract, POPULATION = population, 'MEDIAN INCOME' = median_income)
```


```{r}
burglaries_tracts <- merge(burglaries_tracts, census, by = 'TRACTCE')
```

